Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-04-27 Ben_Manuscripts/thesis/hdphmm.tex
spell checking and mod to usage of subdiffusion

\documentclass{article}

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath} % or simply amstext
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}
\newcommand{\angstrom}{\textup{\AA}}
\newcommand{\colormap}{jet}  % colorbar to use
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{gensymb}
\usepackage{float}
\usepackage{xr}
\usepackage{setspace}
\doublespacing
%\DeclareUnicodeCharacter{03C9}{$\omega$}
\externaldocument[S-]{Supporting_Information}

\title{Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series 
       of Solute Trajectories in Nanostructured Membranes.}

\author{Benjamin J. Coscia, Christopher P. Calderon, Michael R. Shirts}

\begin{document}

  \graphicspath{{./figures/}}
  \maketitle

@@

%  There is a need for highly selective membranes in order to perform efficient 
%  separations of components of complex aqueous streams.
%  %MRS2: add some more examples (When you get around to it).
%  \begin{itemize}
%    \item Organic micropollutants
%    \item Desalination and boric acid removal from seawater.
%    \item While many researchers focus on membrane permeability, we may be 
%    able to reduce costs of commercial nanofiltration and reverse osmosis with
%    higher selectivity.~\cite{werber_materials_2016}
%  \end{itemize}
%
%%  \noindent Amphiphilic molecules are capable of self-assembling into ordered nanostructures.
%%  \begin{itemize}
%%    \item 
%%  \end{itemize}
%
%  Lyotropic liquid crystals (LLC) are a class of amphiphilic molecules whose ordered phases
%  can be cross-linked into mechanically strong membranes capable of highly selective
%  separations.
%  \begin{itemize}
%    \item The shape of the LLC monomers and water content dictates the ordered phase 
%    that they form. There are two phases of particular interest for membrane applications.
%  	\item H\textsubscript{II} phase lyotropic liquid crystals are characterized by 
%  	hexagonally packed, straight, pores while the Q\textsubscript{I} phase consists of
%  	a tortuous network of 3D interconnected pores. 
%  	\item In both cases the pores are uniform in size with radii on the order of 1 nm
%    %MRS2: is it really a molecular weight cutoff?
%    %BJC2: is molecular size better? Obviously there are other factors like shape, but I don't want to get into that
%  	giving them a very strict molecular size cut-off.
%	\item Additionally, they have the potential to disrupt conventional membrane separation
%	techniques by being selective based not only on size and charge, but on chemical
%	functionality as well.
%	\item Their pores are lined with LLC monomer functional groups which can potentially
%	be designed to interact with solutes in a chemically-specific manner
%  \end{itemize}
%
%  There are limits to what we can learn from experiment about LLC membrane design.
%  \begin{itemize}
%    \item Experimental observables like permeability and selectivity allow us to speculate
%    about the molecular origins of separation processes.
%    %MRS2: not quite sure what you mean below?
%    \item This drives an empirical design approach which can potentially neglect key
%    interactions which influence selectivity.
%    \item LLC membranes have been shown to exhibit selectivities which cannot be fully 
%    explained by relatively simple macroscopic models. 
%  \end{itemize}
%  
%  Molecular Dynamics (MD) simulations can give us mechanistic insights with atomistic 
%  resolution so that we can intelligently design new membranes for solute-specific 
%  separations.
%  \begin{itemize}
%    \item In our previous work, we built a detailed atomistic model which we used
%    to understand the nanoscopic structure of an LLC Membrane.~\cite{coscia_understanding_2019}
%    \item We also used the model in order to gain a qualitative understanding of 
%    trapping mechanisms which lead to subdiffusive transport behavior.~\cite{coscia_chemically_2019}
%  \end{itemize}
%
%  Unfortunately, the timescales that we can simulate with MD are insufficient to be
%  able to make well-converged predictions of macroscopic transport properties 
%  traditionally used to characterize membranes in the lab.
%  \begin{itemize}
%    %MRS2: would be good to get this next point into the first line of a paragraph, since it's quite important. 
%    \item However, if we use descriptive stochastic models that can capture solute
%    dynamics, then we could project long timescale behavior in addition to gaining
%    a deeper understanding of solute behavior on short timescales.
%  \end{itemize}

  %%%%%%%%%%%%%%%%%%%%%%%% Thesis chapter starts here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

  In this work we attempt to minimize human effort by applying the infinite hidden 
  Markov Model (IHMM)~\cite{beal_infinite_2002,calderon_inferring_2015}, a nonparameteric 
  Bayesian classification technique, in order to automatically detect and infer the 
  parameters of an unknown number of latent autoregressive (AR) modes present in solute
  center-of-mass time series trajectories.
  We aim to address two primary scientific questions with this work. First, we want
  to know if the IHMM algorithm can help researchers efficiently uncover underlying
  transport mechanisms which give rise to different solute dynamical behavior.
  Second, we want to evaluate the ability of our model to reproduce solute trajectories
  in a stochastic manner so that we can use it to predict macroscopic transport 
  properties which would otherwise require much longer MD simulations.

@@

  We ran all MD simulations and energy minimizations using GROMACS 2018.
  ~\cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}  
  Our python implementation of the IHMM algorithm is available online at 
  \texttt{https://github.com/bencoscia/hdphmm}. All other post-simulation 
  trajectory analysis tools are available online at \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.

@@

  %Based on the algorithms developed by Fox et al.~\cite{fox_sticky_2007}, 
  Using the IHMM framework, we estimated the most likely number and sequence of hidden
  states in each solute center-of-mass trajectory while simultaneously inferring VAR(1)
  parameters for each state and the overall state transition probability matrix, $T$.
  We created a python implementation of this process which we heavily adapted from
  the MATLAB code of Fox et al.~\cite{fox_sticky_2007} Parameter estimation is iterative. 
  Therefore, we looked for convergence of each entry in the $\mathbf{c}$ vector as well
  as the $A$ and $\Sigma$ matrices of Equation~\ref{eqn:var}. We detected equilibration 
  of the parameters using the module \texttt{pymbar.timeseries.detect\_equilibration}~\cite{chodera_simple_2016} on
  the time series of parameter estimates. Plots illustrating parameter convergence are 
  available in Section~\ref{S-section:convergence} of the Supporting Information. We refer the interested reader to 
  much more extensive descriptions of the inference and sampling procedures used to 
  estimate the VAR(1) parameters and the state sequence. ~\cite{beal_infinite_2002,teh_hierarchical_2006,van_gael_beam_2008,fox_nonparametric_2009,fox_bayesian_2010}

  We carefully applied the IHMM algorithm in a way which takes advantage of the
  system's cylindrical symmetry. There are a number of viable ways in which one 
  could choose to analyze these time series. We could use all 3 dimensions straight
  from the output trajectory. We could use just the $z$ dimensional coordinate 
  since we are primarily interested in through-plane membrane transport. Ultimately, 
  we chose to use a multi-step procedure that we believe adequately distinguishes 
  every type of distinct dynamical behavior exhibited by the solutes. The final
  parameterization of the model is in cylindrical coordinates since each pore is a
  cylinder where we expect solutes to exhibit radially symmetric dynamics. We 
  present a step-by-step demonstration of our procedure in Section~\ref{S-section:ihmm_procedure}
  of the Supporting Information.

@@

  We applied the IHMM to each of the 24 solute trajectories independently.
  Although the IHMM is capable of identifying an infinite number of states, 
  a Dirichlet process tends to exhibit a ``rich get richer" effect, favoring
  a fewer number of states.~\cite{dreyer_discovering_2011} By applying the algorithm to each trajectory 
  independently, we reduce the possibility of lumping together multiple 
  similar states which we would prefer to stay separated before clustering.
  Note that the states identified by the IHMM are heavily influenced by 
  the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. 
  In Section~\ref{S-section:prior_guesses} of the Supporting Information we outline our method for 
  choosing reliable prior parameters. The entries of $A$ and $\Sigma$ do 
  not vary over a wide range, so the final parameters were relatively insensitive
  to their priors.

  We ran 2000 iterations of the IHMM procedure in order to arrive at converged 
  state sequences. We evaluated the convergence of the state sequence based on
  the convergence of the VAR parameters. The parameters of each states
  generally converge within 500--1000 iterations. However, the boundaries of 
  each state segment fluctuate which can lead to high variance in the $A$
  and $\Sigma$ parameters in segments with a relatively low number of emissions.
  In these cases, the mean in each dimension tends to be a more reliable indicator of
  convergence (see Figure~\ref{S-fig:convergence3d} of the Supporting Information).

  We defined the finalized parameters of each state as the average of the 
  parameters from each iteration recorded after the equilibration time point.
  We found the equilibration time point of each component of the $\mathbf{c}$ 
  vector as well as $A$ and $\Sigma$ matrices, and then used the longest 
  equilibration time of all dimensions as the equilibration time point.
%  Further details are available in Section~\ref{S-section:convergence} of
%  the Supporting Information.

  We reparameterized the time series, preserving the state sequence, in terms
  of the radial and axial coordinates ($r$, $z$) because $x$ and $y$ aren't 
  meaningful alone due to the system's radial symmetry. We converted the $x$ and 
  $y$ center-of-mass coordinates to $r$. We fixed the state sequence from the 
  3D parameterization and applied the inference component of the IHMM procedure
  to the cylindrical trajectories in order to estimate the VAR(1) parameters in
  terms of $r$ and $z$. Since we fixed the state sequence, and we are using conjugate
  priors, the VAR parameters converge very quickly (see Figure~\ref{S-fig:fixed_state_convergence}).
  Therefore, we only ran the inference procedure for 100 iterations. 
  %However, note that the uncertainty in their
  %estimates are a function of the number of data points used in their estimate.

@@

  We reduced the parameter space used agglomerative clustering, a hierarchical
  clustering approach which uses a linkage criteria in order to successively merge
  similar clusters until a desired intracluster distance threshold or number of
  clusters is reached.~\cite{pedregosa_scikit-learn_2011} We used the Ward linkage criteria, which works to minimize 
  the sum of the squared differences within all clusters.~\cite{ward_hierarchical_1963}
    %MRS6: explain why Ward (make the cluster as minimally disperse?)
   	%BJC6: I'll add some plots to the SI and then write a better justification for
	% the paper. This is lower priority in mind for the thesis.
  We elected to choose the number of clusters rather than the distance threshold.
  For our data, non-parametric methods such as Bayesian Gaussian mixture models~\cite{pedregosa_scikit-learn_2011,gelman_bayesian_2013} tend
  to delocalize the clusters in parameter space (see Section~\ref{S-section:agglomerative}
  of the Supporting Information).

@@

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines. We used the silhouette 
  test in order to score the quality of clustering as a function of the number 
  of clusters chosen.~\cite{kaufman_finding_2009} For our data, the silhouette test generally favors the 
  lowest number of clusters possible (see Section~\ref{S-section:nclusters} of
  the Supporting Information). However, choosing too few clusters tends to 
  not distinguish between visually obvious differences in 
  dynamic behavior. This results in finalized parameter sets that are averages of
  distinct behavior which presents further problems with the predictive modeling
  that we discuss later on. We aimed to maintain the highest silhouette score,
  and thus lowest number of clusters, possible while verifying that visually 
  distinct states stayed separated. Based on these guidelines, we decided to 
  group the $A$ and $\Sigma$ parameters into five clusters each and the radial
  means into three clusters. Of the 75 distinct states allowed by this formulation,
  the solutes in this study showed behavior from 32--36 distinct clusters. While
  this may seem like a high number of states, many are sampled infrequently.

@@

  \subsubsection*{The Dynamics of Methanol}

@@

  It is clear that there is a wealth of information in these plots, but they don't give
  the complete picture needed in order to formulate LLC membrane design principles. One
  can analyze each segment sequentially and form a clear picture of this instance of 
  methanol's behavior without the need to visualize the trajectory with software like VMD.~\cite{humphrey_vmd:_1996}
  However, it is still possible to gain a more general understanding of methanol's average
  behavior.

@@

  %BJC4: I didn't put sodium ion association in this figure because it's infrequent. 
  %BJC4: If I compare to acetic acid or urea, I could add it back. I was thinking of splitting
  % the pi chart in half with slices corresponding to hbonds on one side and association on the other
  \begin{figure}
  \centering
  \begin{subfigure}{0.54\textwidth}
  \includegraphics[width=\textwidth]{hbond_pi_charts.pdf}
  \caption{}\label{fig:hbond_pi_charts}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{hbond_lifetimes_density_MET.pdf}
  \caption{}\label{fig:hbond_lifetimes_density_MET}
  \end{subfigure}
  \caption{(a) Methanol hydrogen bonds in different segments of the LLC monomer
  dependent on the state. States with mean radial locations in the pore region tend to
  hydrogen bond with the head groups and ether linkages. States with mean radial
  locations in the tails tend to hydrogen bond with the tail oxygen atoms.
  (b) Hydrogen bond lifetimes and the local density experienced by methanol
  molecules can help us to understand dynamical states in terms of membrane 
  interactions. The bars are colored according to match the states in Figures~\ref{fig:mechanism_map}
  and~\ref{fig:common_states_MET}. States 1 and 2, which have large covariances, also have short
  hydrogen bond lifetimes and exist in relatively low density regions. The
  hydrogen bond lifetimes of methanol in states with radial means in the tail
  region tend to increase with local density and mean dwell times. 
  %MRS7: not clear from graphic which ones are in the pore.
  %MRS7: is there any legend for the color for b? If same as above figure, indicate. Don't REALLY need color since they bars are already labeled, so it's 2 sources of duplicate information.
  %BJC7: The state numbers are all you need to compare to the previous figure BUT the color make it easier 
  %plot of the methanol trajectory where states aren't labeled.
  %MRS7: cross-hatching doesn't appear in the legend? Not sure which is which (I assume hatched is number density because closer to that axis).
  %BJC7: Is it because the hatches aren't dense enough? I see them in the legend
  This monotonic
  relationship does not exist for states in the pore region. The high 
  local density experienced by state 5 methanol molecules may explain why 
  hydrogen bond lifetimes are short but dwell times are high. State 5 methanol's
  preference towards hydrogen bonds with ether linkages suggests that the solute
  sits between LLC monomer head group which prevent it from wandering to far
  before recovering its previous hydrogen bond.
  }\label{fig:hbond_pichart}
  \end{figure}

@@

  \subsubsection*{Behavior of Other Solutes}

@@

  %BJC5: something we can do for the supporting information to show that this approach is valid: Eliminate nonidealities in our MD trajectories (compared to a pure VAR process) and use toy data to generate trajectories based on some number of states with different parameters. Generate an MSD with those. See if we can reproduce with our model. Can also use this to test new strategies.
  %MRS6: toy data is actually a VERY good idea.  Use a VAR process that looks like MD, and then try to cluster it, and figure out why different.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_MET.pdf}
  \caption{}\label{fig:msd_MET}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_GCL.pdf}
  \caption{}\label{fig:msd_GCL}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_URE.pdf}
  \caption{}\label{fig:msd_URE}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_ACH.pdf}
  \caption{}\label{fig:msd_ACH}
  \end{subfigure}
  \caption{The MSDs predicted based on 1000 realizations of our model (blue) tend to underpredict
  the MSDs based on the MD trajectories (black). Due to the variability in solute behavior across MD trajectories,
  as demonstrated in Figure~\ref{fig:trajectory_realizations_MET}, there are several 
  individual solute MSD curves that are significantly higher than the rest and drive up the
  mean of the MSD. If we remove those trajectories from the MSD calculation, we obtain 
  estimates much closer to the IHMM estimate.(I haven't done it yet, but could plot a third curve
  with outlier removed). Acetic acid, has no obvious outliers which is consistent with our model's
  prediction which lies within the 1$\sigma$ confidence interval of MD after a 1000 ns time lag.}\label{fig:msds}
  \end{figure}

@@

  % BJC6: I'll weaken this for the thesis, but this is what I hope to say in the 
  % final paper.
  Finally, we show how one can generate stochastic realizations of our model that
  can qualitatively and quantitatively reproduce the behavior of MD solute 
  trajectories. The realizations show the hopping and trapping behavior that is
  characteristic of polar solutes in this system. We can also use them to reproduce
  the solute MSDs measured from the MD trajectories. The low computational expense 
  of generating these stochastic trajectories allows one to project solute behavior
  on much longer timescales and thus predict experimentally-relevant macroscopic 
  transport properties like solute flux and selectivity.

  We showcase this modeling approach by example, but it is important to
  recognize the generality of this analysis, especially in the context of molecular
  simulations. Vector autogregressive models can describe a diverse set of behavior
  and so the IHMM may be suited to study many types motion, including those not 
  characterized by hopping and trapping. The IHMM is a powerful approach for 
  understanding particle motion as well as for generating cheap models
  which can give macroscopic insight.

  \section*{Supporting Information}

  Detailed explanations and expansions upon the results and procedures mentioned in
  the main text are described in the Supporting Information. 
  %This information is
  %available free of charge via the Internet at http://pubs.acs.org.

  \section*{Acknowledgments}

  This work was supported in part by the ACS Petroleum Research Fund
  grant \#59814-ND7 and the Graduate Assistance in Areas of National Need (GAANN) 
  fellowship which is funded by the U.S. Department of Education. 
  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.

  \clearpage

  \bibliographystyle{ieeetr}
  \bibliography{hdphmm}

  %\newpage

  %\section*{TOC Graphic}

\end{document}

% LocalWords:  BJC micropollutants Desalination permeability nanofiltration LLC
% LocalWords:  Amphiphilic nanostructures Lyotropic amphiphilic lyotropic MSDDM
% LocalWords:  solutes selectivities solute nanoscopic timescales MSDs MSD HMMs
% LocalWords:  nonparameteric Parrinello Rahman barostat rescale transitioning
% LocalWords:  al's HMM Dirichlet HDP DP Wishart TODO SI al MATLAB teh gael pdf
% LocalWords:  nonparametric bayesian solute's reparameterized intracluster ca
% LocalWords:  delocalize Luzar Chandler histogrammed interpolator methanol's
% LocalWords:  nanostructure carboxylate VMD nonidealities hbonds outlier GAANN
% LocalWords:  autogregressive PRF ACS XSEDE ACI

>>>>

\chapter{Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores
  with Time Series Modeling}

@@

@@

  In this work we attempt to minimize human effort by applying the infinite hidden 
  Markov Model (IHMM)~\cite{beal_infinite_2002,calderon_inferring_2015}, a nonparametric 
  Bayesian classification technique, in order to automatically detect and infer the 
  parameters of an unknown number of latent autoregressive (AR) modes present in solute
  center-of-mass time series trajectories.
  We aim to address two primary scientific questions with this work. First, we want
  to know if the IHMM algorithm can help researchers efficiently uncover underlying
  transport mechanisms which give rise to different solute dynamical behavior.
  Second, we want to evaluate the ability of our model to reproduce solute trajectories
  in a stochastic manner so that we can use it to predict macroscopic transport 
  properties which would otherwise require much longer MD simulations.

@@

  We ran all MD simulations and energy minimizations using GROMACS 2018.
  ~\cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}  
  Our python implementation of the IHMM algorithm is available online at \\
  \texttt{https://github.com/bencoscia/hdphmm}. All other post-simulation 
  trajectory analysis tools are available online at \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.

@@

  %Based on the algorithms developed by Fox et al.~\cite{fox_sticky_2007}, 
  Using the IHMM framework, we estimated the most likely number and sequence of hidden
  states in each solute center-of-mass trajectory while simultaneously inferring VAR(1)
  parameters for each state and the overall state transition probability matrix, $T$.
  We created a python implementation of this process which we heavily adapted from
  the MATLAB code of Fox et al.~\cite{fox_sticky_2007} Parameter estimation is iterative. 
  Therefore, we looked for convergence of each entry in the $\mathbf{c}$ vector as well
  as the $A$ and $\Sigma$ matrices of Equation~\ref{eqn:var}. We detected equilibration 
  of the parameters using the module \texttt{pymbar.timeseries.detect\_equilibration}~\cite{chodera_simple_2016} on
  the time series of parameter estimates. Plots illustrating parameter convergence are 
  available in Section~\ref{section:convergence} of Appendix D. We refer the interested reader to 
  much more extensive descriptions of the inference and sampling procedures used to 
  estimate the VAR(1) parameters and the state sequence. ~\cite{beal_infinite_2002,teh_hierarchical_2006,van_gael_beam_2008,fox_nonparametric_2009,fox_bayesian_2010}

  We carefully applied the IHMM algorithm in a way which takes advantage of the
  system's cylindrical symmetry. There are a number of viable ways in which one 
  could choose to analyze these time series. We could use all 3 dimensions straight
  from the output trajectory. We could use just the $z$ dimensional coordinate 
  since we are primarily interested in through-plane membrane transport. Ultimately, 
  we chose to use a multi-step procedure that we believe adequately distinguishes 
  every type of distinct dynamical behavior exhibited by the solutes. The final
  parameterization of the model is in cylindrical coordinates since each pore is a
  cylinder where we expect solutes to exhibit radially symmetric dynamics. We 
  present a step-by-step demonstration of our procedure in Section~\ref{section:ihmm_procedure}
  of Appendix D.

@@

  We applied the IHMM to each of the 24 solute trajectories independently.
  Although the IHMM is capable of identifying an infinite number of states, 
  a Dirichlet process tends to exhibit a ``rich get richer" effect, favoring
  a fewer number of states.~\cite{dreyer_discovering_2011} By applying the algorithm to each trajectory 
  independently, we reduce the possibility of lumping together multiple 
  similar states which we would prefer to stay separated before clustering.
  Note that the states identified by the IHMM are heavily influenced by 
  the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. 
  In Section~\ref{section:prior_guesses} of Appendix D we outline our method for 
  choosing reliable prior parameters. The entries of $A$ and $\Sigma$ do 
  not vary over a wide range, so the final parameters were relatively insensitive
  to their priors.

  We ran 2000 iterations of the IHMM procedure in order to arrive at converged 
  state sequences. We evaluated the convergence of the state sequence based on
  the convergence of the VAR parameters. The parameters of each states
  generally converge within 500--1000 iterations. However, the boundaries of 
  each state segment fluctuate which can lead to high variance in the $A$
  and $\Sigma$ parameters in segments with a relatively low number of emissions.
  In these cases, the mean in each dimension tends to be a more reliable indicator of
  convergence (see Figure~\ref{fig:convergence3d} of Appendix D).

  We defined the finalized parameters of each state as the average of the 
  parameters from each iteration recorded after the equilibration time point.
  We found the equilibration time point of each component of the $\mathbf{c}$ 
  vector as well as $A$ and $\Sigma$ matrices, and then used the longest 
  equilibration time of all dimensions as the equilibration time point.
%  Further details are available in Section~\ref{section:convergence} of
%  Appendix D.

  We reparameterized the time series, preserving the state sequence, in terms
  of the radial and axial coordinates ($r$, $z$) because $x$ and $y$ aren't 
  meaningful alone due to the system's radial symmetry. We converted the $x$ and 
  $y$ center-of-mass coordinates to $r$. We fixed the state sequence from the 
  3D parameterization and applied the inference component of the IHMM procedure
  to the cylindrical trajectories in order to estimate the VAR(1) parameters in
  terms of $r$ and $z$. Since we fixed the state sequence, and we are using conjugate
  priors, the VAR parameters converge very quickly (see Figure~\ref{fig:fixed_state_convergence}).
  Therefore, we only ran the inference procedure for 100 iterations. 
  %However, note that the uncertainty in their
  %estimates are a function of the number of data points used in their estimate.

@@

  We reduced the parameter space used agglomerative clustering, a hierarchical
  clustering approach which uses a linkage criteria in order to successively merge
  similar clusters until a desired intra cluster distance threshold or number of
  clusters is reached.~\cite{pedregosa_scikit-learn_2011} We used the Ward linkage criteria, which works to minimize 
  the sum of the squared differences within all clusters.~\cite{ward_hierarchical_1963}
    %MRS6: explain why Ward (make the cluster as minimally disperse?)
   	%BJC6: I'll add some plots to the SI and then write a better justification for
	% the paper. This is lower priority in mind for the thesis.
  We elected to choose the number of clusters rather than the distance threshold.
  For our data, non-parametric methods such as Bayesian Gaussian mixture models~\cite{pedregosa_scikit-learn_2011,gelman_bayesian_2013} tend
  to delocalize the clusters in parameter space (see Section~\ref{section:agglomerative}
  of Appendix D).

@@

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines. We used the silhouette 
  test in order to score the quality of clustering as a function of the number 
  of clusters chosen.~\cite{kaufman_finding_2009} For our data, the silhouette test generally favors the 
  lowest number of clusters possible (see Section~\ref{section:nclusters} of
  Appendix D). However, choosing too few clusters tends to 
  not distinguish between visually obvious differences in 
  dynamic behavior. This results in finalized parameter sets that are averages of
  distinct behavior which presents further problems with the predictive modeling
  that we discuss later on. We aimed to maintain the highest silhouette score,
  and thus lowest number of clusters, possible while verifying that visually 
  distinct states stayed separated. Based on these guidelines, we decided to 
  group the $A$ and $\Sigma$ parameters into five clusters each and the radial
  means into three clusters. Of the 75 distinct states allowed by this formulation,
  the solutes in this study showed behavior from 32--36 distinct clusters. While
  this may seem like a high number of states, many are sampled infrequently.

@@

  \subsubsection{The Dynamics of Methanol}

@@

  It is clear that there is a wealth of information in these plots, but they don't give
  the complete picture needed in order to formulate LLC membrane design principles. One
  can analyze each segment sequentially and form a clear picture of this instance of 
  methanol's behavior without the need to visualize the trajectory with software like VMD.%~\cite{humphrey_vmd:_1996}
  However, it is still possible to gain a more general understanding of methanol's average
  behavior.

@@

  %BJC4: I didn't put sodium ion association in this figure because it's infrequent. 
  %BJC4: If I compare to acetic acid or urea, I could add it back. I was thinking of splitting
  % the pi chart in half with slices corresponding to hbonds on one side and association on the other
  \begin{figure}
  \centering
  \begin{subfigure}{0.54\textwidth}
  \includegraphics[width=\textwidth]{hbond_pi_charts.pdf}
  \caption{}\label{fig:hbond_pi_charts}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{hbond_lifetimes_density_MET.pdf}
  \caption{}\label{fig:hbond_lifetimes_density_MET}
  \end{subfigure}
  \caption{(a) Methanol hydrogen bonds in different segments of the LLC monomer
  dependent on the state. States with mean radial locations in the pore region tend to
  hydrogen bond with the head groups and ether linkages. States with mean radial
  locations in the tails tend to hydrogen bond with the tail oxygen atoms.
  (b) Hydrogen bond lifetimes and the local density experienced by methanol
  molecules can help us to understand dynamical states in terms of membrane 
  interactions. The bars are colored according to match the states in Figures~\ref{fig:mechanism_map}
  and~\ref{fig:common_states_MET}. States 1 and 2, which have large covariances, also have short
  hydrogen bond lifetimes and exist in relatively low density regions. The
  hydrogen bond lifetimes of methanol in states with radial means in the tail
  region tend to increase with local density and mean dwell times. 
  %MRS7: not clear from graphic which ones are in the pore.
  %MRS7: is there any legend for the color for b? If same as above figure, indicate. Don't REALLY need color since they bars are already labeled, so it's 2 sources of duplicate information.
  %BJC7: The state numbers are all you need to compare to the previous figure BUT the color make it easier 
  %plot of the methanol trajectory where states aren't labeled.
  %MRS7: croshatching doesn't appear in the legend? Not sure which is which (I assume hatched is number density because closer to that axis).
  %BJC7: Is it because the hatches aren't dense enough? I see them in the legend
  This monotonic
  relationship does not exist for states in the pore region. The high 
  local density experienced by state 5 methanol molecules may explain why 
  hydrogen bond lifetimes are short but dwell times are high. State 5 methanol's
  preference towards hydrogen bonds with ether linkages suggests that the solute
  sits between LLC monomer head group which prevent it from wandering to far
  before recovering its previous hydrogen bond.
  }\label{fig:hbond_pichart}
  \end{figure}

@@

  \subsubsection{Behavior of Other Solutes}

@@

  %BJC5: something we can do for Appendix D to show that this approach is valid: Eliminate nonidealities in our MD trajectories (compared to a pure VAR process) and use toy data to generate trajectories based on some number of states with different parameters. Generate an MSD with those. See if we can reproduce with our model. Can also use this to test new strategies.
  %MRS6: toy data is actually a VERY good idea.  Use a VAR process that looks like MD, and then try to cluster it, and figure out why different.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_MET.pdf}
  \caption{}\label{fig:msd_MET}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_GCL.pdf}
  \caption{}\label{fig:msd_GCL}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_URE.pdf}
  \caption{}\label{fig:msd_URE}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_ACH.pdf}
  \caption{}\label{fig:msd_ACH}
  \end{subfigure}
  \caption{The MSDs predicted based on 1000 realizations of our model (blue) tend to underpredict
  the MSDs based on the MD trajectories (black). Due to the variability in solute behavior across MD trajectories,
  as demonstrated in Figure~\ref{fig:trajectory_realizations_MET}, there are several 
  individual solute MSD curves that are significantly higher than the rest and drive up the
  mean of the MSD. If we remove those trajectories from the MSD calculation, we obtain 
  estimates much closer to the IHMM estimate. Acetic acid, has no obvious outliers which is
  consistent with our model's prediction which lies within the 1$\sigma$ confidence interval of MD 
  after a 1000 ns time lag.}\label{fig:msds}
  \end{figure}

@@

  Finally, we show how one can generate stochastic realizations of our model that
  can qualitatively and quantitatively reproduce the behavior of MD solute 
  trajectories. The realizations show the hopping and trapping behavior that is
  characteristic of polar solutes in this system. We can also use them to reproduce
  the solute MSDs measured from the MD trajectories. The low computational expense 
  of generating these stochastic trajectories allows one to project solute behavior
  on much longer timescales and thus predict experimentally-relevant macroscopic 
  transport properties like solute flux and selectivity.

  We showcase this modeling approach by example, but it is important to
  recognize the generality of this analysis, especially in the context of molecular
  simulations. Vector autoregressive models can describe a diverse set of behavior
  and so the IHMM may be suited to study many types motion, including those not 
  characterized by hopping and trapping. The IHMM is a powerful approach for 
  understanding particle motion as well as for generating cheap models
  which can give macroscopic insight.
